Clinical prediction with localized modeling using similarity-based cohorts: A scoping review

### 2023-08-18 docs/Manuscript/Manuscript.tex
update zotero keys + make tables consistent

{[}25{]} & Prediction\hspace{6em} & Practice\hspace{6em} &

@@

{[}26{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &

@@

{[}30{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &

@@

{[}31{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &

@@

Clinical\hspace{6em} & Longitudinal\hspace{6em} & 245,825 & 1,466,474 \\
{[}34{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &

@@

{[}35{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
Clinical\hspace{6em} & Longitudinal\hspace{6em} & 245,825 & 1,466,474 \\

@@

{\centering \includegraphics[width=0.8\linewidth]{../../fig/fig-years}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1218}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4365}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4416}}@{}}
\caption{\label{tab:composite}Methodological elements of studies
included in the synthesis.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Elements
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Terminology
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Elements
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Terminology
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
{[}11{]} & predictive modeling on similarity cohorts & CBR; targeted
subsampling \\
{[}19{]} & supervised similarity learning & CBR; case-structured
retrieval \\
{[}20{]} & unsupervised similarity learning & Combined Kohonen type
neural network-case-based reasoning \\
{[}21{]} & supervised similarity learning & Transductive inference \\
{[}22{]} & unsupervised similarity learning & CBR; entropic distance
measure \\
{[}23{]} & tiered constraint; similarity matching & CBR \\
{[}24{]} & interactive implementation & CBR; eXiT*CBR \\
{[}27{]} & supervised similarity learning & CBR \\
{[}28{]} & supervised similarity learning & CBR \\
{[}25{]} & predictive modeling on similarity cohorts & Local modeling;
personalized modeling; TWNFI \\
{[}26{]} & predictive modeling on similarity cohorts & Personalized
modeling; TWNFI \\
{[}29{]} & supervised similarity learning; predictive modeling on
similarity cohorts & CBR \\
{[}30{]} & unclear & CBR \\
{[}31{]} & whole-population weight learning & Gaussian processes \\
{[}32{]} & predictive modeling on similarity cohorts & Personalized
model \\
{[}34{]} & descriptive modeling on similarity cohorts & CBR \\
{[}33,35{]} & supervised similarity learning; predictive modeling on
similarity cohorts & Similarity model, personalized model, precision
cohort, personalized treatment options \\
{[}13{]} & supervised cohort construction & CBR, statistical case-based
reasoning \\
{[}12{]} & predictive modeling on similarity cohorts & K-nearest
neighbors survival \\
{[}14{]} & predictive modeling on similarity cohorts & Personalized
prediction \\
{[}15{]} & predictive modeling on similarity cohorts; explanatory
modeling on similarity cohorts & Personalized predictive model \\
{[}16{]} & predictive modeling on similarity cohorts & Patient-specific
predictive model \\
{[}17{]} & predictive modeling on similarity cohorts & Personalized
predictive modeling \\
{[}18{]} & predictive modeling on similarity cohorts & Personalized
model \\
\end{longtable}

@@

Toward detection and diagnosis, {[}20{]} proposed a hybrid neural
net--CBR system to distinguish (with confidence bounds) arthritic versus
control patients, based on several histological features. {[}28{]} used
collaborative multilabel CBR to subtype melanoma patients based on
confocal and dermoscopy images. {[}17{]} used localized models built
from a multi-type additive similarity measure to distinguish type 2
diabetic versus control populations. Along the way, {[}21{]}, {[}18{]},
and {[}26{]} took an iterative model-building approach to several tasks:
predicting glomerular filtration rate, a key indicator of renal
function, from demographic and physiological variables; identifying
patients with colon cancer using a large number of gene expression
measurements; and identifying patients with type 2 diabetes based on
demographic, physiological, and genetic variables.

While several studies emphasized the potential or actual value to
decision-making of their methods and tools, only one incorporated
treatment decisions into their approach: By taking ``decision points''
as their units of analysis, {[}33{]} and {[}35{]} built not classifiers
or predictors but comparative effectiveness models into a localized
framework, providing for the first time in our sample explicitly
prescriptive rather than descriptive clinical decision support.

@@

CBR can be understood as the opposite side of a trade-off with
rule-based reasoning between model size and model complexity; {[}36{]}
describe their approach as ``knowledge-light'', in that ``the cases do
not contain explicit explanation structures; instead, explanation is
achieved by comparison of the query case with retrieved cases''. This
means that the greatest performance and efficiency challenges in CBR
have to do with the retrieval and revision phases in the R4 cycle.
Several studies addressed these challenges: {[}13{]}, while excluded
from the synthesis, were the earliest to propose that cohorts be bounded
by a similarity threshold rather than by a number of cases, and this
improved performance in their experiments. {[}38{]} leveraged predictor
weights obtained from logistic regression to inform the similarity
calculation used in retrieval. {[}12{]} used non-parametric models to
reduce the computational burden of revision (prediction). {[}32{]}
proposed to improve efficiency along the entire R4 cycle, but in
particular by only executing task-dependent steps in real time
(``just-in-time learning''), and {[}35{]} and {[}33{]} partitioned
multiple phases in the R4 cycle into offline and online components, only
the latter of which would be performed in real time as new data are
received.

Other studies addressed limitations of available tools. {[}24{]}
implemented a comprehensive CBR tool in response to the lack of
general-purpose software, to enable coupling with other tools as well as
to expedite development, experimentation, and uptake. Several other
teams also set out to develop more interoperable clinical support tools
{[}15,22,25,31{]}.

@@

The motivational and methodological unity of these studies does not
reflect a unified research program. Besides the lack of any primary
journal of record, we observed collaborations only among the authors of
smaller contiguous programs, including applications of the TWNFI
methodology {[}21,26{]}, individualized mortality prediction for ICU
patients {[}14,16{]} and the use of more explanatory models to
prioritize predictors or treatments for chronic disease {[}15,33,35{]}.

These programs used significantly varying terminology for common
concepts, and in particular no common term is in use for what we here
refer to as localized modeling; authors described this kind of method as
``targeted prognosis'' {[}11{]}, ``transductive inference'' {[}21{]},
``personalized decision support'' {[}14{]}, ``personalized (predictive)
models'' {[}25, in contrast to ``local models,''15,17,32{]}, and
``precision cohort'' analysis {[}17,33,35{]}. We find uses of
``targeted'', ``personalized'', and ``precision'' generic and imprecise
to this approach, while transductive inference is an established term
for a broader set of methods in ML. Because local naturally contrasts
with global, we propose ``localized models'' as a suitable term for this
counterpart to globally-fitted models.

@@

The 25 studies included in our synthesis used composite techniques that
we organized into two major types. One type, similarity learning, was
used in 8 studies that tie in to the much larger literature on metric
learning {[}39{]}. These techniques are designed to detect structure in
data, especially associations between predictors and responses, and to
use this structure to inform the definition of a patient similarity
measure. Seven studies used response values in training data sets to
supervise similarity learning while two used unsupervised learning
(Table \ref{tab:composite}). Based on the key properties of metric
learning algorithms identified by {[}40{]}, most learned measures were
non-linear while some were local, and some were optimized globally while
others locally. However, not all qualified as metric learning, since the
resulting measures would not necessarily satisfy the triangle
inequality. Each then used its similarity measure to retrieve training
cases relevant to each testing case. Once retrieved, most took a
standard NN approach to generating predictions from the cohort; two
exceptions {[}29,33{]} proceeded to fit predictive models to the
retrieved cohorts and are discussed below.

@@

In addition to {[}12{]}, {[}25{]}, {[}15{]}, {[}16{]}, {[}29{]},
{[}33{]}, and {[}35{]}, five other studies employed localized models:
{[}11{]}, {[}14{]}, {[}26{]}, {[}17{]}, and {[}32{]}. In most cases
these models were purely predictive in application; of the exceptions,
{[}11{]} produced purely descriptive models of survival outcomes, which
were evaluated for their precision rather than their accuracy, while
{[}15{]} used generalized regression models descriptively as well as
predictively, as did some of the same authors {[}33,35{]} later. The
most common approach to cohort selection was to optimize a size
threshold via manual exploration {[}11,14--17,29{]} or via
cross-validation {[}12,26{]}. Further in the former direction, one study
{[}32{]} set a fixed cohort size while two {[}25,33,35{]} devised more
sophisticated algorithms to balance multiple desiderata including
predictive accuracy.

Several themes emerged from this sample. Foremost was the optimization
of retrieved cohort sizes for some measure of model performance. Despite
the early recommendation by {[}13{]} to threshold cohorts by similarity
rather than by cardinality, only the one previously published study
{[}11{]} and the most recent study {[}33,35{]} in this corpus took this
approach. Viewed as a hyperparameter of the individualized modeling
approach, cohort size can be treated in the same way as neighborhood
size in NN prediction (viewed here as a special case of localized
modeling wherein the model is a simple summary statistic), so we will
not discuss it further.

An alternative approach was to optimize multiple hyperparameters
together: {[}18{]} proposed an iterative algorithm to settle on an
optimal number and set of predictors as well as number of training cases
(comprising the retrieved cohort) that was later used by {[}25{]}. More
recently, {[}33{]} and {[}35{]} proposed a three-step process for cohort
selection that filtered by exact match for one subset of predictors,
used domain-informed similarity measures to rank these, and optimized
the similarity threshold for a trade-off between cohort size and a
measure of bias called cohort balance. Notably, though their similarity
measure was supervised, this optimization process was not.

With the exception of the descriptive survival models of {[}11{]}, only
one thread in this corpus concerned itself with the interpretability of
localized models: {[}15{]} fit logistic regression models to
similarity-based cohorts and examined not only their predictive
performance but the localized sets of largest and most detectable
predictors, termed risk profiles, and how they differed across the
testing population. This approach informed their later use of localized
comparative effectiveness--style models to recommend courses of
treatment at decision points during a monitored patent's stay
{[}33,35{]}. These studies suggest much wider potential for localized
real-world evidence generation, which has long been promoted as a
promise of advances in clinical and health informatics {[}1{]}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998a}{}}%
\CSLLeftMargin{{[}5{]} }%
\CSLRightInline{Gierl L, Bull M, Schmidt R. {CBR} in {Medicine}. In:
Lenz M, Burkhard H-D, Bartsch-Spörl B, Wess S, editors. Case-{Based
Reasoning Technology}, vol. 1400, {Berlin, Heidelberg}: {Springer Berlin
Heidelberg}; 1998, p. 273--97.
\url{https://doi.org/10.1007/3-540-69351-3_11}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Mariuzzi1997}{}}%
\CSLLeftMargin{{[}11{]} }%
\CSLRightInline{Mariuzzi G, Mombello A, Mariuzzi L, Hamilton PW, Weber
JE, Thompson D, et al. Quantitative study of ductal breast
cancer--patient targeted prognosis: An exploration of case base
reasoning. Pathology, Research and Practice 1997;193:535--42.
\url{https://doi.org/10.1016/S0344-0338(97)80011-8}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Yearwood1997a}{}}%
\CSLLeftMargin{{[}19{]} }%
\CSLRightInline{Yearwood J, Wilkinson R. Retrieving cases for treatment
advice in nursing using text representation and structured text
retrieval. Artif Intell Med 1997;9:79--99.
\url{https://doi.org/10.1016/s0933-3657(96)00362-4}.}

\leavevmode\vadjust pre{\hypertarget{ref-Wyns2004}{}}%
\CSLLeftMargin{{[}20{]} }%
\CSLRightInline{Wyns B, Sette S, Boullart L, Baeten D, Hoffman IEA, De
Keyser F. Prediction of diagnosis in patients with early arthritis using
a combined {Kohonen} mapping and instance-based evaluation criterion.
Artif Intell Med 2004;31:45--55.
\url{https://doi.org/10.1016/j.artmed.2004.01.002}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Elter2007}{}}%
\CSLLeftMargin{{[}22{]} }%
\CSLRightInline{Elter M, Schulz-Wendtland R, Wittenberg T. The
prediction of breast cancer biopsy outcomes using two {CAD} approaches
that both emphasize an intelligible decision process. Med Phys
2007;34:4164--72. \url{https://doi.org/10.1118/1.2786864}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Lopez2011}{}}%
\CSLLeftMargin{{[}24{]} }%
\CSLRightInline{López B, Pous C, Gay P, Pla A, Sanz J, Brunet J.
{eXiT}*{CBR}: {A} framework for case-based medical diagnosis development
and experimentation. Artif Intell Med 2011;51:81--91.
\url{https://doi.org/10.1016/j.artmed.2010.09.002}.}

\leavevmode\vadjust pre{\hypertarget{ref-Liang2015}{}}%
\CSLLeftMargin{{[}25{]} }%
\CSLRightInline{Liang W, Hu Y, Kasabov N. Evolving personalized modeling
system for integrated feature, neighborhood and parameter optimization
utilizing gravitational search algorithm. EVOLVING SYSTEMS 2015;6:1--14.
\url{https://doi.org/10.1007/s12530-013-9081-x}.}

\leavevmode\vadjust pre{\hypertarget{ref-Verma2015a}{}}%
\CSLLeftMargin{{[}26{]} }%
\CSLRightInline{Verma A, Fiasche M, Cuzzola M, Irrera G. Comparative
study of existing personalized approaches for identifying important gene
markers and for risk estimation in {Type2 Diabetes} in {Italian}
population. EVOLVING SYSTEMS 2015;6:15--22.
\url{https://doi.org/10.1007/s12530-013-9083-8}.}

\leavevmode\vadjust pre{\hypertarget{ref-CampilloGimenez2012a}{}}%
\CSLLeftMargin{{[}27{]} }%
\CSLRightInline{Campillo-Gimenez B, Bayat S, Cuggia M. Coupling
{K-nearest} neighbors with logistic regression in case-based reasoning.
Stud Health Technol Inform 2012;180:275--9.}

\leavevmode\vadjust pre{\hypertarget{ref-Nicolas2014a}{}}%
\CSLLeftMargin{{[}28{]} }%
\CSLRightInline{Nicolas R, Fornells A, Golobardes E, Corral G, Puig S,
Malvehy J. {DERMA}: A melanoma diagnosis platform based on collaborative
multilabel analog reasoning. ScientificWorldJournal 2014;2014:351518.
\url{https://doi.org/10.1155/2014/351518}.}

\leavevmode\vadjust pre{\hypertarget{ref-Vilhena2016a}{}}%
\CSLLeftMargin{{[}29{]} }%
\CSLRightInline{Vilhena J, Vicente H, Martins MR, Grañeda JM, Caldeira
F, Gusmão R, et al. A case-based reasoning view of thrombophilia risk. J
Biomed Inform 2016;62:265--75.
\url{https://doi.org/10.1016/j.jbi.2016.07.013}.}

\leavevmode\vadjust pre{\hypertarget{ref-Malykh2018a}{}}%
\CSLLeftMargin{{[}30{]} }%
\CSLRightInline{Malykh VL, Rudetskiy SV. Approaches to {Medical
Decision-Making Based} on {Big Clinical Data}. J Healthc Eng
2018;2018:3917659. \url{https://doi.org/10.1155/2018/3917659}.}

\leavevmode\vadjust pre{\hypertarget{ref-Zhang2018a}{}}%
\CSLLeftMargin{{[}31{]} }%
\CSLRightInline{Zhang H, Zhu F, Dodge H, Higgins G, Omenn G, Guan Y, et
al. A similarity-based approach to leverage multi-cohort medical data on
the diagnosis and prognosis of {Alzheimer}'s disease. GIGASCIENCE
2018;7. \url{https://doi.org/10.1093/gigascience/giy085}.}

\leavevmode\vadjust pre{\hypertarget{ref-Ma2020a}{}}%
\CSLLeftMargin{{[}32{]} }%
\CSLRightInline{Ma X, Si Y, Wang Z, Wang Y. Length of stay prediction
for {ICU} patients using individualized single classification algorithm.
COMPUTER METHODS AND PROGRAMS IN BIOMEDICINE 2020;186.
\url{https://doi.org/10.1016/j.cmpb.2019.105224}.}

\leavevmode\vadjust pre{\hypertarget{ref-Tang2021d}{}}%
\CSLLeftMargin{{[}33{]} }%
\CSLRightInline{Tang P, Miller S, Stavropoulos H, Kartoun U, Zambrano J,
Ng K. Precision population analytics: Population management at the
point-of-care. JOURNAL OF THE AMERICAN MEDICAL INFORMATICS ASSOCIATION
2021;28:588--95. \url{https://doi.org/10.1093/jamia/ocaa247}.}

>>>>

{[}25{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &
{[}26{]} & Prediction\hspace{6em} & Practice\hspace{6em} &

@@

@@

{[}30{]} & Diagnosis\hspace{6em} & Knowledge\hspace{6em} &
{[}31{]} & Decision Support\hspace{6em} & Practice\hspace{6em} &

@@

@@

@@

{[}34{]}; {[}35{]} & Decision Support\hspace{6em} & Practice\hspace{6em}
& Clinical\hspace{6em} & Longitudinal\hspace{6em} & 245,825 &
1,466,474 \\

@@

{\centering \includegraphics[width=0.75\linewidth]{../../fig/fig-years}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1128}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4410}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4462}}@{}}
\caption{\label{tab:composite}Methodological elements of studies
included in the synthesis.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Elements
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Terminology
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Elements
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Terminology
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
{[}19{]} & supervised similarity learning & CBR; case-structured
retrieval \\
{[}11{]} & predictive modeling on similarity cohorts & CBR; targeted
subsampling \\
{[}20{]} & unsupervised similarity learning & Combined Kohonen type
neural network-case-based reasoning \\
{[}13{]} & supervised cohort construction & CBR, statistical case-based
reasoning \\
{[}21{]} & supervised similarity learning & Transductive inference \\
{[}22{]} & unsupervised similarity learning & CBR; entropic distance
measure \\
{[}23{]} & tiered constraint; similarity matching & CBR \\
{[}24{]} & interactive implementation & CBR; eXiT*CBR \\
{[}18{]} & predictive modeling on similarity cohorts & Personalized
model \\
{[}25{]} & predictive modeling on similarity cohorts & Personalized
modeling; TWNFI \\
{[}26{]} & predictive modeling on similarity cohorts & Local modeling;
personalized modeling; TWNFI \\
{[}12{]} & predictive modeling on similarity cohorts & K-nearest
neighbors survival \\
{[}27{]} & supervised similarity learning & CBR \\
{[}28{]} & supervised similarity learning & CBR \\
{[}15{]} & predictive modeling on similarity cohorts; explanatory
modeling on similarity cohorts & Personalized predictive model \\
{[}14{]} & predictive modeling on similarity cohorts & Personalized
prediction \\
{[}29{]} & supervised similarity learning; predictive modeling on
similarity cohorts & CBR \\
{[}16{]} & predictive modeling on similarity cohorts & Patient-specific
predictive model \\
{[}30{]} & whole-population weight learning & Gaussian processes \\
{[}31{]} & unclear & CBR \\
{[}32{]} & predictive modeling on similarity cohorts & Personalized
model \\
{[}17{]} & predictive modeling on similarity cohorts & Personalized
predictive modeling \\
{[}33{]} & descriptive modeling on similarity cohorts & CBR \\
{[}34{]}; {[}35{]} & supervised similarity learning; predictive modeling
on similarity cohorts & Similarity model, personalized model, precision
cohort, personalized treatment options \\
\end{longtable}

@@

Toward detection and diagnosis, {[}20{]} proposed a hybrid neural
net--CBR system to distinguish (with confidence bounds) arthritic versus
control patients, based on several histological features. {[}28{]} used
collaborative multilabel CBR to subtype melanoma patients based on
confocal and dermoscopy images. {[}17{]} used localized models built
from a multi-type additive similarity measure to distinguish type 2
diabetic versus control populations. Along the way, {[}21{]}, {[}18{]},
and {[}25{]} took an iterative model-building approach to several tasks:
predicting glomerular filtration rate, a key indicator of renal
function, from demographic and physiological variables; identifying
patients with colon cancer using a large number of gene expression
measurements; and identifying patients with type 2 diabetes based on
demographic, physiological, and genetic variables.

While several studies emphasized the potential or actual value to
decision-making of their methods and tools, only one incorporated
treatment decisions into their approach: By taking ``decision points''
as their units of analysis, {[}34{]} and {[}35{]} built not classifiers
or predictors but comparative effectiveness models into a localized
framework, providing for the first time in our sample explicitly
prescriptive rather than descriptive clinical decision support.

@@

CBR can be understood as the opposite side of a trade-off with
rule-based reasoning between model size and model complexity; {[}36{]}
describe their approach as ``knowledge-light'', in that ``the cases do
not contain explicit explanation structures; instead, explanation is
achieved by comparison of the query case with retrieved cases''. This
means that the greatest performance and efficiency challenges in CBR
have to do with the retrieval and revision phases in the R4 cycle.
Several studies addressed these challenges: {[}13{]}, while excluded
from the synthesis, were the earliest to propose that cohorts be bounded
by a similarity threshold rather than by a number of cases, and this
improved performance in their experiments. {[}38{]} leveraged predictor
weights obtained from logistic regression to inform the similarity
calculation used in retrieval. {[}12{]} used non-parametric models to
reduce the computational burden of revision (prediction). {[}32{]}
proposed to improve efficiency along the entire R4 cycle, but in
particular by only executing task-dependent steps in real time
(``just-in-time learning''), and {[}35{]} and {[}34{]} partitioned
multiple phases in the R4 cycle into offline and online components, only
the latter of which would be performed in real time as new data are
received.

Other studies addressed limitations of available tools. {[}24{]}
implemented a comprehensive CBR tool in response to the lack of
general-purpose software, to enable coupling with other tools as well as
to expedite development, experimentation, and uptake. Several other
teams also set out to develop more interoperable clinical support tools
{[}15,22,26,30{]}.

@@

The motivational and methodological unity of these studies does not
reflect a unified research program. Besides the lack of any primary
journal of record, we observed collaborations only among the authors of
smaller contiguous programs, including applications of the TWNFI
methodology {[}21,25{]}, individualized mortality prediction for ICU
patients {[}14,16{]} and the use of more explanatory models to
prioritize predictors or treatments for chronic disease {[}15,34,35{]}.

These programs used significantly varying terminology for common
concepts, and in particular no common term is in use for what we here
refer to as localized modeling; authors described this kind of method as
``targeted prognosis'' {[}11{]}, ``transductive inference'' {[}21{]},
``personalized decision support'' {[}14{]}, ``personalized (predictive)
models'' {[}26, in contrast to ``local models,''15,17,32{]}, and
``precision cohort'' analysis {[}17,34,35{]}. We find uses of
``targeted'', ``personalized'', and ``precision'' generic and imprecise
to this approach, while transductive inference is an established term
for a broader set of methods in ML. Because local naturally contrasts
with global, we propose ``localized models'' as a suitable term for this
counterpart to globally-fitted models.

@@

The 25 studies included in our synthesis used composite techniques that
we organized into two major types. One type, similarity learning, was
used in 8 studies that tie in to the much larger literature on metric
learning {[}39{]}. These techniques are designed to detect structure in
data, especially associations between predictors and responses, and to
use this structure to inform the definition of a patient similarity
measure. Seven studies used response values in training data sets to
supervise similarity learning while two used unsupervised learning
(Table \ref{tab:composite}). Based on the key properties of metric
learning algorithms identified by {[}40{]}, most learned measures were
non-linear while some were local, and some were optimized globally while
others locally. However, not all qualified as metric learning, since the
resulting measures would not necessarily satisfy the triangle
inequality. Each then used its similarity measure to retrieve training
cases relevant to each testing case. Once retrieved, most took a
standard NN approach to generating predictions from the cohort; two
exceptions {[}29,34{]} proceeded to fit predictive models to the
retrieved cohorts and are discussed below.

@@

In addition to {[}12{]}, {[}26{]}, {[}15{]}, {[}16{]}, {[}29{]},
{[}34{]}, and {[}35{]}, five other studies employed localized models:
{[}11{]}, {[}14{]}, {[}25{]}, {[}17{]}, and {[}32{]}. In most cases
these models were purely predictive in application; of the exceptions,
{[}11{]} produced purely descriptive models of survival outcomes, which
were evaluated for their precision rather than their accuracy, while
{[}15{]} used generalized regression models descriptively as well as
predictively, as did some of the same authors {[}34,35{]} later. The
most common approach to cohort selection was to optimize a size
threshold via manual exploration {[}11,14--17,29{]} or via
cross-validation {[}12,25{]}. Further in the former direction, one study
{[}32{]} set a fixed cohort size while two {[}26,34,35{]} devised more
sophisticated algorithms to balance multiple desiderata including
predictive accuracy.

Several themes emerged from this sample. Foremost was the optimization
of retrieved cohort sizes for some measure of model performance. Despite
the early recommendation by {[}13{]} to threshold cohorts by similarity
rather than by cardinality, only the one previously published study
{[}11{]} and the most recent study {[}34,35{]} in this corpus took this
approach. Viewed as a hyperparameter of the individualized modeling
approach, cohort size can be treated in the same way as neighborhood
size in NN prediction (viewed here as a special case of localized
modeling wherein the model is a simple summary statistic), so we will
not discuss it further.

An alternative approach was to optimize multiple hyperparameters
together: {[}18{]} proposed an iterative algorithm to settle on an
optimal number and set of predictors as well as number of training cases
(comprising the retrieved cohort) that was later used by {[}26{]}. More
recently, {[}34{]} and {[}35{]} proposed a three-step process for cohort
selection that filtered by exact match for one subset of predictors,
used domain-informed similarity measures to rank these, and optimized
the similarity threshold for a trade-off between cohort size and a
measure of bias called cohort balance. Notably, though their similarity
measure was supervised, this optimization process was not.

With the exception of the descriptive survival models of {[}11{]}, only
one thread in this corpus concerned itself with the interpretability of
localized models: {[}15{]} fit logistic regression models to
similarity-based cohorts and examined not only their predictive
performance but the localized sets of largest and most detectable
predictors, termed risk profiles, and how they differed across the
testing population. This approach informed their later use of localized
comparative effectiveness--style models to recommend courses of
treatment at decision points during a monitored patent's stay
{[}34,35{]}. These studies suggest much wider potential for localized
real-world evidence generation, which has long been promoted as a
promise of advances in clinical and health informatics {[}1{]}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998}{}}%
\CSLLeftMargin{{[}5{]} }%
\CSLRightInline{Gierl L, Bull M, Schmidt R. {CBR} in {Medicine}. In:
Lenz M, Burkhard H-D, Bartsch-Spörl B, Wess S, editors. Case-{Based
Reasoning Technology}, vol. 1400, {Berlin, Heidelberg}: {Springer Berlin
Heidelberg}; 1998, p. 273--97.
\url{https://doi.org/10.1007/3-540-69351-3_11}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Mariuzzi1997}{}}%
\CSLLeftMargin{{[}11{]} }%
\CSLRightInline{Mariuzzi G, Mombello A, Mariuzzi L, Hamilton PW, Weber
JE, Thompson D, et al. Quantitative study of ductal breast
cancer--patient targeted prognosis: An exploration of case base
reasoning. Pathology - Research and Practice 1997;193:535--42.
\url{https://doi.org/10.1016/S0344-0338(97)80011-8}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Yearwood1997a}{}}%
\CSLLeftMargin{{[}19{]} }%
\CSLRightInline{Yearwood J, Wilkinson R. Retrieving cases for treatment
advice in nursing using text representation and structured text
retrieval. Artificial Intelligence in Medicine 1997;9:79--99.
\url{https://doi.org/10.1016/s0933-3657(96)00362-4}.}

\leavevmode\vadjust pre{\hypertarget{ref-Wyns2004a}{}}%
\CSLLeftMargin{{[}20{]} }%
\CSLRightInline{Wyns B, Sette S, Boullart L, Baeten D, Hoffman IEA, De
Keyser F. Prediction of diagnosis in patients with early arthritis using
a combined {Kohonen} mapping and instance-based evaluation criterion.
Artificial Intelligence in Medicine 2004;31:45--55.
\url{https://doi.org/10.1016/j.artmed.2004.01.002}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Elter2007a}{}}%
\CSLLeftMargin{{[}22{]} }%
\CSLRightInline{Elter M, Schulz-Wendtland R, Wittenberg T. The
prediction of breast cancer biopsy outcomes using two {CAD} approaches
that both emphasize an intelligible decision process. Medical Physics
2007;34:4164--72. \url{https://doi.org/10.1118/1.2786864}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Lopez2011a}{}}%
\CSLLeftMargin{{[}24{]} }%
\CSLRightInline{López B, Pous C, Gay P, Pla A, Sanz J, Brunet J.
{eXiT}*{CBR}: {A} framework for case-based medical diagnosis development
and experimentation. Artificial Intelligence in Medicine 2011;51:81--91.
\url{https://doi.org/10.1016/j.artmed.2010.09.002}.}

\leavevmode\vadjust pre{\hypertarget{ref-Verma2015a}{}}%
\CSLLeftMargin{{[}25{]} }%
\CSLRightInline{Verma A, Fiasche M, Cuzzola M, Irrera G. Comparative
study of existing personalized approaches for identifying important gene
markers and for risk estimation in {Type2 Diabetes} in {Italian}
population. Evolving Systems 2015;6:15--22.
\url{https://doi.org/10.1007/s12530-013-9083-8}.}

\leavevmode\vadjust pre{\hypertarget{ref-Liang2015a}{}}%
\CSLLeftMargin{{[}26{]} }%
\CSLRightInline{Liang W, Hu Y, Kasabov N. Evolving personalized modeling
system for integrated feature, neighborhood and parameter optimization
utilizing gravitational search algorithm. Evolving Systems 2015;6:1--14.
\url{https://doi.org/10.1007/s12530-013-9081-x}.}

\leavevmode\vadjust pre{\hypertarget{ref-CampilloGimenez2012a}{}}%
\CSLLeftMargin{{[}27{]} }%
\CSLRightInline{Campillo-Gimenez B, Bayat S, Cuggia M. Coupling
{K-nearest} neighbors with logistic regression in case-based reasoning.
Studies in Health Technology and Informatics 2012;180:275--9.}

\leavevmode\vadjust pre{\hypertarget{ref-Nicolas2014a}{}}%
\CSLLeftMargin{{[}28{]} }%
\CSLRightInline{Nicolas R, Fornells A, Golobardes E, Corral G, Puig S,
Malvehy J. {DERMA}: A melanoma diagnosis platform based on collaborative
multilabel analog reasoning. The Scientific World Journal
2014;2014:351518. \url{https://doi.org/10.1155/2014/351518}.}

\leavevmode\vadjust pre{\hypertarget{ref-Vilhena2016a}{}}%
\CSLLeftMargin{{[}29{]} }%
\CSLRightInline{Vilhena J, Vicente H, Martins MR, Grañeda JM, Caldeira
F, Gusmão R, et al. A case-based reasoning view of thrombophilia risk.
Journal of Biomedical Informatics 2016;62:265--75.
\url{https://doi.org/10.1016/j.jbi.2016.07.013}.}

\leavevmode\vadjust pre{\hypertarget{ref-Zhang2018a}{}}%
\CSLLeftMargin{{[}30{]} }%
\CSLRightInline{Zhang H, Zhu F, Dodge H, Higgins G, Omenn G, Guan Y, et
al. A similarity-based approach to leverage multi-cohort medical data on
the diagnosis and prognosis of {Alzheimer}'s disease. GigaScience
2018;7. \url{https://doi.org/10.1093/gigascience/giy085}.}

\leavevmode\vadjust pre{\hypertarget{ref-Malykh2018a}{}}%
\CSLLeftMargin{{[}31{]} }%
\CSLRightInline{Malykh VL, Rudetskiy SV. Approaches to {Medical
Decision-Making Based} on {Big Clinical Data}. Journal of Healthcare
Engineering 2018;2018:3917659.
\url{https://doi.org/10.1155/2018/3917659}.}

\leavevmode\vadjust pre{\hypertarget{ref-Ma2020a}{}}%
\CSLLeftMargin{{[}32{]} }%
\CSLRightInline{Ma X, Si Y, Wang Z, Wang Y. Length of stay prediction
for {ICU} patients using individualized single classification algorithm.
Computer Methods and Programs in Biomedicine 2020;186.
\url{https://doi.org/10.1016/j.cmpb.2019.105224}.}

\leavevmode\vadjust pre{\hypertarget{ref-Wang2020}{}}%
\CSLLeftMargin{{[}33{]} }%
\CSLRightInline{Wang Y, Liang Y, Sun H, Yang Y. Emergency {Response} for
{COVID-19 Prevention} and {Control} in {Urban Rail Transit Based} on
{Case-Based Reasoning Method}. Discrete Dynamics in Nature \& Society
2020:1--9.}

\leavevmode\vadjust pre{\hypertarget{ref-Tang2021c}{}}%
\CSLLeftMargin{{[}34{]} }%
\CSLRightInline{Tang P, Miller S, Stavropoulos H, Kartoun U, Zambrano J,
Ng K. Precision population analytics: Population management at the
point-of-care. Journal of the American Medical Informatics Association
2021;28:588--95. \url{https://doi.org/10.1093/jamia/ocaa247}.}

\leavevmode\vadjust pre{\hypertarget{ref-Ng2021b}{}}%
\CSLLeftMargin{{[}35{]} }%
\CSLRightInline{Ng K, Kartoun U, Stavropoulos H, Zambrano J, Tang P.
Personalized treatment options for chronic diseases using precision
cohort analytics. Scientific Reports 2021;11.
\url{https://doi.org/10.1038/s41598-021-80967-5}.}
